---
title: "Homework 6"
output: pdf_document
date: "2025-11-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r prob1}
# Defines parameters
shape <- 256000
rate <- 16000
n <- 34
x_bar <- 16.01

# Calculates mu and sigma parameters
mu <- shape / rate
sigma <- sqrt(shape) / rate

# Calculates standard error (se)
se <- sigma / sqrt(n)

# Calculates z-score
z <- (x_bar - mu) / se

# Calculates probability that the sample mean (x_bar) > 16.01
probability <- pnorm(z, lower.tail = FALSE)
probability

```
The probability that the average fill of the sampled soda cans is greater than 16.01 ounces is ```r probability```.




## Problem 2

```{r prob2}
# Below is the code for parts a - c (and the explanation for part d) of problem 2
# Defines parameters
sigma <- 1
sample <- 60
```


a)
```{r prob2a}
# Calculates mean and standard deviation
mu <- 8.2
mu
sd <- sigma / sqrt(sample)
sd

```
The mean of the sampling distribution of the sample mean is ```r mu``` and the standard deviation is ```r sd```.


b)
```{r prob2b}
# Calculates 90th percentile for sample mean
prob_90_q2 <- qnorm(0.90, mean = mu, sd = sd)
prob_90_q2

```
Interpretation: The value obtained from part b implies that 90% of possible random users from the sample will have a tablet usage time of ```r prob_90_q2``` minutes or less.


c)
```{r prob2c}
# Calculates + or - 1 standard deviation
p1 <- pnorm(1) - pnorm(-1)
# Calculates + or - 2 standard deviation
p2 <- pnorm(2) - pnorm(-2)
# Calculates + or - 3 standard deviation
p3 <- pnorm(3) - pnorm(-3)
c(p1, p2, p3)
```
$\pm \mathbf{1}$ Standard Deviation: ```r p1```

$\pm \mathbf{2}$ Standard Deviation: ```r p2```

$\pm \mathbf{3}$ Standard Deviation: ```r p3```


d) We can solve part c by using the Empirical Rule (the 68-95-99.7% rule): A basic generalization is that $\pm \mathbf{1}$ Standard Deviation is $\approx$ 68%, $\pm \mathbf{2}$ Standard Deviation is $\approx$ 95%, and $\pm \mathbf{3}$ Standard Deviation is $\approx$ 99.7%.




## Problem 3

```{r prob3}
# Below is the code for parts a - d of problem 3
# Defines parameters
n <- 75
mu <- 5 * 0.5 
sigma <- sqrt(5 * 0.5 * 0.5)
sd <- sigma / sqrt(n)
x_bar <- 2.25
```


a) 
```{r prob3a}
# Calculates P(x_bar < 2.25)
z <- (x_bar - mu) / sd
prob_avg_score <- pnorm(z)
prob_avg_score
```


b) 
```{r prob3b}
# Calculates 90th percentile
prob_90_q3b <- qnorm(0.90, mean = mu, sd = sd)
prob_90_q3b
```


c) 
```{r prob3c}
z_total_score <- (200 - (n * mu)) / (sqrt(n) * sigma)
prob_total <- pnorm(z_total_score)
prob_total
```


d) 
```{r prob3d}
prob_90_q3d <- qnorm(0.90, mean = n * mu, sd = sqrt(n) * sigma)
prob_90_q3d
```




## Problem 4

```{r prob4}
```


a) 
```{r prob4a}
# Defines parameters 
lambda <- 0.5
x <- 2.5

# Calculates probability of data usage being > 2.5
pexp(x, rate = lambda, lower.tail = FALSE)

```


b) 
```{r prob4b}
mu <- 2
sigma <- 2 / sqrt(80)
x_bar <- 2.5

pnorm(x_bar, mean = mu, sd = sigma, lower.tail = FALSE)
```


c) The probabilities are different because part a follows an exponential distribution and measures the probability of an individual, whereas part b follows a normal distribution (central limit theorem) and measures the probability of a sample mean.


## Problem 5

```{r prob5}
# Defines parameters
p_hat <- 22/30
n <- 30
N <- 70

# Calculates standard error
se <- sqrt(p_hat * (1 - p_hat) / n) * sqrt((N - n) / (N - 1))

# Calculates confidence interval
ci_q5 <- p_hat + c(-1, 1) * qnorm(0.975) * se
ci_q5
```
The 95% confidence interval is (```r ci_q5```) and 65% falls under that range.



## Problem 6

```{r prob6}
# Define parameters
mu <- 180
sigma <- 25
n <- 15
max_weight_per_person <- 3500

# Calculate total mean and standard deviation
mu_total <- n * mu
sd_total <- sqrt(n) * sigma

# Calculates z score and probablity of total weight > 3500
z <- (max_weight_per_person - mu_total) / sd_total
prob_exceeds_weight <- pnorm(z, lower.tail = FALSE)
prob_exceeds_weight

```




## Problem 7

```{r prob7}
# Defines parameters
n <- 30
mu_nightly <- 25
var_nightly <- 25

# Calculates total mean and variance values
mu_total <- n * mu_nightly
var_total <- sqrt(n * var_nightly)

# Calculates z score
z <- (600 - mu_total) / var_total

# Calculates probability
prob_bottles_sold <- pnorm(z, lower.tail = FALSE)
prob_bottles_sold

```





## Problem 8

```{r prob8}
# Defines parameters
p_hat <- 2/30
n <- 30
N <- 54

# Calculates standard error
se <- sqrt(p_hat * (1 - p_hat) / n) * sqrt((N - n) / (N - 1))

# Calculates confidence interval
ci_q8 <- p_hat + c(-1, 1) * qnorm(0.975) * se
ci_q8

```
The 95% confidence interval is (```r ci_q8```) and 13.1% does not fall in that range as it's greater, which suggests that the class has a lower rate of left-handed people than the overall US population.


## Problem 9

```{r prob9}
# Loads dataset
library("UsingR")
data(babies)

# Ensures the ages are within a valid range
valid_ages <- complete.cases(babies$age, babies$dage)

# Performing the t test on the ages
t.test(babies$dage[valid_ages], babies$age[valid_ages], paired = TRUE, conf.level = 0.95)

```
Interpretation: The confidence interval does not contain 0. The data implies that on average, fathers are a few years older than mothers, but there is no significant difference in ages.